02-antecedentes

### 2024-06-28 02-antecedentes.qmd

Despite high-income inequality and limited social mobility in Chile, and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023]. The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context, as reflected in surveys conducted by the Center for Public Studies (CEP). According to this data, 35.9% prefer private health insurance, and 63% would prefer private education. Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs (Centro de Estudios Públicos, 2024).

@@

Figure @fig-hypotheses depicts a summary of the research hypothesis described before. Market justice preferences are our main concept under study (vis-a-vis dependent variable), which refers to the justification of better access to health, education, and pensions based on payment capacities (as we detail further in the methods section). The explanatory concepts (independent variables) are at two levels: student (individual) variables and school (context) variables. The moderation hypotheses are represented through arrows pointing to the arrow between meritocratic perception and market justice preferences, which will be estimated through interaction effects.

>>>>

Despite high-income inequality and limited social mobility in Chile, and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023]. The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context, as reflected in surveys conducted by the Center for Public Studies (CEP). According to this data, 35.9% prefer private health insurance, and 63% would prefer private education. Yet, research on children's justification in this area remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs [@centrodeestudiospublicos_estudio_2024].

@@

@fig-hypotheses depicts a summary of the research hypothesis described before. Market justice preferences are our main concept under study (vis-a-vis dependent variable), which refers to the justification of better access to health, education, and pensions based on payment capacities (as we detail further in the methods section). The explanatory concepts (independent variables) are at two levels: student (individual) variables and school (context) variables. The moderation hypotheses are represented through arrows pointing to the arrow between meritocratic perception and market justice preferences, which will be estimated through interaction effects.
